### 2016-05-16 ms/ms_probaweb.tex
Edits from all

% TODO:
% Add Ben and Spencer's edits
% Revise the model with fixed interaction pribabilites (havens)
% Compile the references
% Distribute for a last round of edits
% Add scussion on other on other types of interactions

@@

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommy Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$, Spencer A. Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

@@

4: \\

@@

6:\\

7:\\

8:\\

@@

11:\\

@@

\textbf{Figures:}

\textbf{Tables:}

@@

Biogeography has historically focused on the spatial distribution and abundance
of species. Both are driven by the way species interact with one another,
but also by the way these interactions vary across time and space. Here,
we call for an integrated approach in which community structure represented
as a network of ecological interactions, and show how it translates to
biogeography questions. Under this representation, the niche is intergratively
defined to be both the effect of the environment on species distribution (the
Grinnellian dimension of the niche) and on the ecological interactions among
them (the Eltonian dimension). Starting from this novel concept, we develop a
quantitative theory to explain turnover of interactions in space and time -–
\emph{i.e.} a novel approach to interaction distribution modelling. We apply
this framework to host–parasite interactions across Europe and find that
two aspects of the environment (temperature and precipitation) leave a strong
imprint on species co-occurrence, but not on species interactions. Even where
species co-occur, interaction proves a stochastic rather than deterministic
process, adding to variation in realized network structure. We also find
that a large majority of species pairs are never found together, thus
precluding any inferences regarding their probability to interact. Our
framework provides a basis to explain the variation of network structure
at large spatial scales and opens new perspectives at the frontier between
species distribution modelling and community ecology.

@@

Because it is an explicit description of species interactions, the network
approach offers a convenient representation of communities. Species are
represented as nodes, so that networks already encompass all the information
used by current approaches; in addition, interactions are represented by links,
so that networks provide additional, higher-order information on community
structure. To date, studies of network diversity have mostly been concerned
with the distribution of interactions within locations, and less so with
the variation among locations \citep{Dunne2005, Bascompte2007, Ings2007,
Kefi2012}. There is however ample evidence that interaction networks vary
in space and time \citep{Laliberte2010, Poisot2012, Albouy2014, Poisot2016,
Trojelsgaard2015}. Metacommunity theory provides explanations for the variation
in the distribution species \citep{Gravel2011c, Pillai2011}, but there is
no such explanation to the variation of node and link occurrences, nor do
we expect their distribution to covary in a positive way. Consequently,
we urgently need a conceptual framework to formalize these observations,
as it is the only way towards fulfilling the goal of community ecology:
providing cogent predictions about, and understanding of, the structure of
ecollogical communities.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a clear need to bring the two together.
Here, we offer an integrated approach to do so, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche that integrates the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a quantitative theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. We apply this novel
framework to a large data set on host–parasite interactions across Europe and
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Figure 1. The regional pool of interactions, or metaweb, represents
potential interactions among all species that could be found in a given area.
In this particular case, there are 275 nodes, and 1173 links among the plants (52
nodes), herbivores (96 nodes), and parasitoids (127 nodes) from Northern
Europe. An instance of a local community is also illustrated, with 45 nodes
and 93 interactions. Only 55.0\% of all potential interactions are realized locally,
revealing the stochastic nature of ecological interactions. Our objective here
is to provide a conceptual framework to explain the sampling of the regional
pool of interactions, along with a quantitative method to predict it. The
problem could be formalized sequentially by first understanding why only a
fraction of the species are co-occurring locally and second why these species
do or do not interact.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either because colonization and extinction events \citep{Hanski1999} are inherently
unpredictable or because of strong non-linear feedbacks in community dynamics
generating alternative transients and equilibria \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. A more recent approach decomposes the total variation of the
community data into species and site contributions to beta diversity
\citep{Legendre2013}, and further partitions into dissimilarity due to changes
in species richness and those due to actual species turnover \citep{Baselga2010;
Carvalho et al. 2012}. Even though these methods compare whole lists of
species among sites or measurements, they remain fundamentally `species-based',
since they report the variation within columns. None of them
explicitly considers the variation of associations (i.e., of pairs or higher-order
motifs – \citealt{Stouffer2007}).

The ``niche'' is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
BIOMOD \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions within this approach
\citep{Peterson2011}. As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

As mentioned above, in the network approach to community structure, species
and interactions are represented by nodes and links, respectively.
Associations can also be represented by matrices in which entries represent
the occurrence or intensity of interactions among species (rows and columns).
Network complexity is then computed as the number of interactions (in the case
of binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition was found to be independent of the variation of ecological interactions,
suggesting that these two components of network variability may well respond to
different drivers \citep{Poisot2012}.

Interestingly, the ecological network literature also has its own ’niche
model’ to position a species in a community \citep{Williams2000}. The niche of a
species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over 3 to 5 different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure and its variants have successfully explained the
complexity of a variety of networks, from food webs to plant–pollinator
systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

A more integrative description of the niche is key to understanding spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far beyond the
``n-dimensional hypervolume'' defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been frequently
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and improve it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}. The
Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. This aspect of the niche is well known
by community ecologists, but is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization \citep{Williams2010; Gravel2013}
made it more operational.

These perspectives are rather orthogonal to each other, and this has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
``[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions''.
Their representation merges zero-net growth isoclines delimiting the
Grinnellian niche (when does the population persists) with impact vectors delimiting
the Eltonian niche (what is the per-capita impact). While this representation
has been very influential in local-scale community ecology (the resource-ratio
theory of coexistence, \citealt{Tilman1982}), it remains impractical at larger
spatial scales because of the difficulties in measuring it. The absence of any
mathematical representation of the niche that could be easily fit to
ecological data may explain why biogeographers are still struggling to develop
species distribution models that also consider ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components (Fig. 2). The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources that allow interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
that allow a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.

@@

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. In particular, we seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define
$L_{ijy}$ as a stochastic variable, and focus on the probability that
this event occurs, $P(L_{ijy})$. We note that the occurrence of an interaction is dependent
on the co-occurrence of species $i$ and $j$. This argument might seem trivial
at first, but the explicit consideration of this condition in the
probabilistic representation of ecological interactions will prove instrumental
to understand their variation. We define $X_{iy}$ as a stochastic variable
representing the occurrence of species $i$ at location $y$. The quantity we seek to
understand is the probability of a joint event, conditional on the set of
environmental conditions $E_y$:

@@

Or simply said, the probability of observing both species $i$ and $j$ plus an
interaction between $i$ and $j$ given the conditions $E_y$. This probability
could be decomposed into two parts using the product rule of probabilities:

@@

The first term on the right-hand side of the equation is the probability
of observing the two species co-occurring at location $y$. It corresponds
to the Grinnellian dimension of the niche. The second term represents the
probability that an interaction occurs between species $i$ and $j$, given that
they are co-occurring. This predicate can be refined using information on
trait distribution and trait matching rules (\textbf{Bartomeus Func Ecol}).
Above, we referred to this entity as the ``metaweb'' and it corresponds to
the Eltonian dimension of the niche. Below, we will see how this formalism
can be directly fit to empirical data. But before turning to an application,
we will discuss the interpretation of different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation, there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
do in fact constrain distributions \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent, and only the marginal occurrence must be accounted for, $P(
|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}). Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is a topic of its own, however, and is influenced by both network
topology and species richness \citep{Cazelles2016}. Not only do direct interactions
influence co-occurrence, but indirect interactions do as well (e.g. plant species
sharing an herbivore could repel each other in space). The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-occurrence
is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2015}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will focus here on
the variation of interactions and not on their distribution, and leave this
specific issue for the Perspectives and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}|X_{ijy}=1) = 1$ and $P(L_{ijy}|X_{ijy}=0) = 0$. This approach might
be a reasonable approximation if the scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation solely arises from species
distributions.

Ecological interactions could also vary with the environment, so that $P(L_{ijy}
|E_y)$. Although it is not common to see a conditional representation of
pairwise ecological interactions, experimental studies have revealed them to
frequently be sensitive to the environment. For instance, \citep{Mckinnon2010} showed
that predation risks of shorebirds vary at the continental scale, decreasing
from the south to the north. It is also common to see increasing top-down
control with temperature (e.g. \citealt{Shurin2012, Gray2016}). Effects of the
environment on interactions also propagate up the community and influence network
structure \citep{Tylianakis2007, Woodward2010; Petchey2010; Lelong2015}.

@@

We now turn to an illustration of our framework with the analysis of an
empirical dataset of host–parasite networks sampled throughout the south–north
environmental gradient in continental Europe. The focal system consists of
local food webs of willows (genus Salix), their galling insects, and the
parasitoids of these gallers. Targeting this system, we ask: i) how much does
network structure vary across the gradient, and ii) what is the primary driver
of network turnover across the gradient?

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}, and is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The parasitoid community includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.

>>>>

@@

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommi Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$, Spencer A. Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason M. Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

@@

4: Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA.\\

@@

6: Ecology and Evolutionary Biology Department, University of Arizona, P.O. Box 210088, Tucson, AZ 85721, USA\\

7: University of Eastern Finland, Department of Environmental and Biological Sciences, P.O. Box 111, FI-80101 Joensuu, Finland\\

8: Université de Montréal, Département des Sciences Biologiques, 90 Avenue Vincent d’Indy, Montréal, QC H2V3S9, Canada.\\

@@

11: Department of Life Sciences, Imperial College London, Silwood Park Campus, 
Buckhurst Road, Ascot, Berkshire SL5 7PY, United Kingdom\\

@@

\textbf{Figures: 6}

\textbf{Tables: 2}

@@

Biogeography has traditionally focused on the spatial distribution and
abundance of species. Both are driven by the way species interact with one
another, but also by the way these interactions vary across time and space.
Here, we call for an integrated approach in which community structure
represented as a network of ecological interactions, and show how it
translates to biogeography questions. We propose that the ecological niche may
be redefined to encompass the effect of the environment on species
distribution (the Grinnellian dimension of the niche) and on the ecological
interactions among them (the Eltonian dimension). Starting from this 
concept, we develop a quantitative theory to explain turnover of interactions
in space and time -– \emph{i.e.} a novel approach to interaction distribution
modelling. We apply this framework to host–parasite interactions across Europe
and find that two aspects of the environment (temperature and precipitation)
exert a strong imprint on species co-occurrence, but not on species
interactions. Even where species co-occur, interaction proves to be stochastic
rather than deterministic, adding to variation in realized network structure.
We also find that a large majority of host-parasite pairs are never found
together, thus precluding any inferences regarding their probability to
interact. This first attempt to explain variation of network
structure at large spatial scales opens new perspectives at the interface
of species distribution modelling and community ecology.

@@

Because it is an explicit description of species interactions, network
approaches offer a convenient representation of communities. Species are
represented as nodes, so that networks already encompass all the information
used by current approaches; in addition, interactions are represented by
links, so that networks provide additional, higher-order information on
community structure. To date, studies of network diversity have mostly been
concerned with the distribution of interactions within locations, and less so
with variation among locations \citep{Dunne2005, Bascompte2007, Ings2007,
Kefi2012}. There is, however, ample evidence that interaction networks vary in
space and time \citep{Laliberte2010, Poisot2012, Albouy2014, Poisot2016,
Trojelsgaard2015}, even though there is no common framework with which to
generalize these results. Metacommunity theory provides explanations for
variation in the distribution of the nodes \citep{Gravel2011c, Pillai2011},
but there is no such explanation to the variation of node and link
occurrences. Consequently, we urgently need a conceptual framework to
formalize these observations, as it is the only way towards fulfilling the
goal of community ecology: providing cogent predictions about, and
understanding of, the structure of ecological communities.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a clear need to bring the two together.
Here, we offer an integrated approach to do so, adopting the view that
community structure is best represented as a network of ecological
interactions. Based on this idea, we propose a new description of the basic
concept of the ecological niche that integrates the effect of the environment
on species distribution and on the ecological interactions among them.
Starting from this redefined concept, we develop a quantitative theory to
explain turnover of interactions in space and time. We first present the
conceptual framework, and then formalize it mathematically, using a
probabilistic model to represent the sampling of the regional pool of
interactions. At the level of species pairs, the statistical approach could be
conceived as an interaction distribution model. At the community level, the
approach provides a likelihood-based method to compare different hypotheses of
network turnover. As an illustrative example, we apply this novel framework to
a large data set on host–parasite interactions across Europe and perform test
hypotheses underlying variation of species and interactions. The network
structure changes systematically across the latitudinal gradient, with a peak
of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, or \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized locally, revealing the stochastic nature of
ecological interactions. Our objective here is to provide a conceptual
framework to explain the sampling of the regional pool of interactions, along
with a quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-
occur locally and second why these species do or do not interact.

There are multiple causes of spatial turnover of species co-occurrence. The first
and most-studied driver is the effect of variation in the abiotic environment
on species performance. Combined with specific responses in demography, it
generates variation among sites by selecting the locally fittest species
\citep{Leibold2004}. Stochasticity plays an additional role, either because
colonization and extinction events \citep{Hanski1999} are inherently
unpredictable or because strong non-linear feedbacks in community dynamics
generate alternative transients and equilibria \citep{Chase2007, Vellend2014}.
Analyses of community turnover are usually performed with data represented in
a table with rows corresponding to sites (or measurements) and columns to
species. Metrics of beta diversity quantify the variance of this community
data \citep{Legendre2005}. Traditional approaches rely on measures of
dissimilarity among communities, such as the Jaccard or Bray–Curtis indices.
More recent approaches decompose total variation of the community data into
species and site contributions to beta diversity \citep{Legendre2013}, and
further partition it into dissimilarity due to changes in species richness and
dissimilarity due to actual species turnover \citep{Baselga2010;
Carvalho2012}. Even though these methods compare whole lists of species among
sites or measurements, they remain fundamentally `species-based', since they
report variation within columns. None of them explicitly considers variation
of associations (i.e., of pairs or higher-order motifs –
\citealt{Stouffer2007}).

Similarly, we are getting a better understanding of interaction turnover. As
mentioned above, in the network approach to community structure, species and
interactions are represented by nodes and links, respectively. Associations
can also be represented by matrices in which entries represent the occurrence
or intensity of interactions among species (rows and columns). Network
complexity is then computed as the number of interactions (in the case of
binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition can be independent of variation of ecological
interactions, suggesting that species and interaction distribution may well
respond to different drivers \citep{Poisot2012}.

The "niche" is by far the dominant concept invoked to explain species
distributions and community assembly, from the local to the global scale.
Following \citealt{Hutchinson1957}, the niche is viewed as the set of
environmental conditions allowing a population to establish and persist (see
also \citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to put into practice with species distribution models, as it
maps naturally on available distributional and environmental data. Consequently, a vast array of statistical tools have been developed to
implement it (e.g. BIOMOD \citealt{Thuiller2003}, MaxEnt
\citealt{Phillips2006}). It is however much harder to account for ecological
interactions within this approach \citep{Peterson2011}. As such, these
interactions are often viewed as externalities constraining or expanding the
range of environmental conditions required for a species to maintain a viable
population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

Interestingly, the ecological network literature also has its own "niche
model" to position a species in a community \citep{Williams2000}. The niche of
a species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure and its variants have successfully
explained the complexity of a variety of networks, from food webs to
plant–pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

Despite several attempts to update the concept of the ecological niche,
ecologists have not moved far beyond the "n-dimensional hypervolume" defined
by Hutchinson. Despite its intuitive interpretation and easy translation into
species distribution models \citep{Boulangeat2012, Blonder2014}, the concept
has been frequently criticized \citep{Hardin1960, Peters1991, Silvertown2004},
and several attempts have been made to expand and improve it
\citep{Pulliam2000, Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}.
The Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. While this aspect of the niche is well
known by community ecologists, it is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization using functional traits
\citep{Gravel2013; Bartomeus2016} made it more operational.

These perspectives are rather orthogonal to each other, and this has resulted
in considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
"[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions".
Their representation merges zero-net-growth isoclines delimiting the
Grinnellian niche ("when does the population persists?"") with impact vectors
delimiting the Eltonian niche ("what is the per-capita impact?"). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}), it remains
impractical at larger spatial scales because of the difficulties in measuring
it. The absence of any mathematical representation of the niche that can be
easily fit to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions. Thus, a more integrative description of the niche
will be key to understand spatial and temporal turnover in community
structure.

We propose to integrate the two perspectives of the niche using a visual
representation of both components (Fig. 2). The underlying rationale is that,
in addition to the environmental constraints on demographic performance (Fig.2
top panel), any organism requires resources to meet its metabolic demands and
to sustain reproduction (Fig. 2 nodes in network of bottom panel). Abiotic
environmental axes are any non-consumable factors affecting the demographic
performance of an organism. Alternatively, the resource axes are traits of the
resources that allow interactions with the consumer. The niche can therefore
be viewed as the set of abiotic environmental variables (the Grinnellian
component) along with the set of traits (the Eltonian component) that allow a
population to establish and to persist at a location. Accordingly, each
species can be characterized by an optimal position along both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.

@@

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. In particular, we seek to represent the
probability that an interaction between species $i$ and $j$ occurs at location
$y$. We define $L_{ijy}$ as a stochastic variable, and focus on the
probability that this event occurs, $P(L_{ijy})$. We note that the occurrence
of an interaction is dependent on the co-occurrence of species $i$ and $j$.
This argument might seem trivial at first, but the explicit consideration of
this condition in the probabilistic representation of ecological interactions
will prove instrumental to understanding their variation. We define $X_{iy}$
as a stochastic variable representing the occurrence of species $i$ at
location $y$. The quantity we seek to understand is the probability of a joint
event, conditional on the set of environmental conditions $E_y$:

@@

Or simply said, the probability of observing both species $i$ and $j$ plus an
interaction between $i$ and $j$ given the conditions $E_y$ at location $y$.
This probability could be decomposed into two parts using the product rule of
probabilities:

@@

The first term on the right-hand side of the equation is the probability
of observing the two species co-occurring at location $y$. It corresponds
to the Grinnellian dimension of the niche. The second term represents the
probability that an interaction occurs between species $i$ and $j$, given that
they are co-occurring. This predicate can be refined using information on
trait distribution and trait matching rules (\citep{Bartomeus2016}).
Above, we referred to this entity as the "metaweb" and it corresponds to
the Eltonian dimension of the niche. Below, we will see how this formalism
can be directly fit to empirical data. But before turning to an application,
we will discuss the interpretation of different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses concerning temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation, there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
constrain distribution \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent, and only the marginal occurrence must be accounted for, $P(
X_{ijy}}|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976,
Connor1980, Gotelli2000}). Second- and higher-order interactions (e.g. trophic
cascades) could also affect co-occurrence. Co-occurrence of multiple species
embedded in ecological networks is a topic of its own, however, and is
influenced by both network topology and species richness \citep{Cazelles2016}.
Not only direct interactions influence co-occurrence, but indirect
interactions do as well (e.g. plant species sharing an herbivore, or
herbivores sharing parasitoids, could repel each other in space
\citealt{Holt1993}). The impact of direct interactions and first-order
indirect interactions on co-occurrence tends to vanish with increasing species
richness in the community. Further, co-occurrence is also influenced by the
covariance of interacting species to an environmental gradient
\citep{Cazelles2015}. Because of the complexity of relating co-occurrence to
the structure of interaction networks, we will focus here on the variation of
interactions and not on their distribution, and leave this specific issue for
the Perspectives section and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, rather than
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}|X_{ijy}=1) = 1$ and $P(L_{ijy}|X_{ijy}=0) = 0$. This approach might
be a reasonable approximation if the spatial or temporal scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation arises solely from species
distributions.

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy} |E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to  be sensitive to the environment. For instance,
\citep{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions also propagate up the
community and influence network structure \citep{Tylianakis2007, Woodward2010;
Petchey2010; Lelong2015}.

@@

We now turn to an illustration of our framework with the analysis of an
empirical dataset of host–parasite networks sampled throughout the south–north
environmental gradient in continental Europe. The focal system consists of
local food webs of willows (genus Salix), their galling insects, and the
natural enemies (parasitoids and inquilines) of these gallers. Targeting this
system, we ask: i) how much does network structure vary across the gradient,
and ii) what is the primary driver of network turnover across the gradient?

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}, and is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.
